### 2022-02-15 doc/src/AdvPLIC.tex
Writes to genmsi are ignored while Busy is true
Fix lots of typos

Given the domain hierarchies depicted in the figures, if it were
necessary to partition the application harts for multiple OSes,
machine-level software would need to prevent direct OS access to the
supervisor-level interrupt domain and instead provide SBI services
for controlling PLIC interrupts or, alternatively, emulate the control
interfaces of separate superivsor-level interrupt domains, one for each
OS.
Note that such emulation might still make use of the PLIC's
physical supervisor-level interrupt domain, but under the control of
machine-level software.
\end{commentary}

@@

If \z{genmsi} is written while Busy is true, the extempore MSI that was
previously pending may be discarded without ever being sent.

@@

A read from \z{claimi} that returns a value of zero has the simulaneous
side effect of setting the \z{iforce} register to zero.

@@

The \z{genmsi} register (Section~\ref{sec:AdvPLIC-reg-genmsi}) exists
to allow software to determinie when all earlier MSIs have arrived at a
hart.
To use \z{genmsi} for this purpose, software can dedicate one external
interrupt identity at each hart's IMSIC interrupt file solely for PLIC
synchronization.
Assuming there are multiple harts, a PLIC's \z{genmsi} register should
also be protected by a standard mutual-exclusion lock.
The following sequence can then be used to synchronize between a PLIC
and a specific hart:
\begin{enumerate}

>>>>

Given the domain hierarchies depicted in the figures, if it were
necessary to partition the application harts for multiple OSes,
machine-level software would need to prevent direct OS access to the
supervisor-level interrupt domain and instead provide SBI services
for controlling PLIC interrupts or, alternatively, emulate the control
interfaces of separate supervisor-level interrupt domains, one for each
OS.
Note that such emulation might still make use of the PLIC's
physical supervisor-level interrupt domain, but under the control of
machine-level software.
\end{commentary}

@@

While Busy is true, writes to \z{genmsi} are ignored.

@@

A read from \z{claimi} that returns a value of zero has the simultaneous
side effect of setting the \z{iforce} register to zero.

@@

The \z{genmsi} register (Section~\ref{sec:AdvPLIC-reg-genmsi}) exists
to allow software to determine when all earlier MSIs have arrived at a
hart.
To use \z{genmsi} for this purpose, software can dedicate one external
interrupt identity at each hart's IMSIC interrupt file solely for PLIC
synchronization.
Assuming there are multiple harts, a PLIC's \z{genmsi} register should
also be protected by a standard mutual-exclusion lock.
The following sequence can then be used to synchronize between a PLIC
and a specific hart:
\begin{enumerate}

### 2022-02-15 doc/src/IMSIC.tex
For IMSICs, address ranges should be reserved for every possible hart group number
Fix lots of typos

The layout of an interupt-file's memory region is:\nopagebreak
\begin{displayLinesTable}[l@{\quad}l@{\qquad}l]
offset    & \ size  & register name \\
\noalign{\medskip}
\z{0x000} & 4 bytes & \z{seteipnum\_le} \\
\z{0x004} & 4 bytes & \z{seteipnum\_be} \\
\end{displayLinesTable}
All other bytes in an interrupt file's 4-KiB memory region are reserved
and must be implemented as read-only zeros.

@@

The requirements for read-only-zero pages apply only within each group.
That is, for each group~$g$, within the address ranges
\begin{displayLinesTable}[lcll]
$g\times\mbox{2}^{E}+A$ & through &
  $g\times\mbox{2}^{E}+A+\mbox{2}^{k+C}-\mbox{1}$, &
  and\\
$g\times\mbox{2}^{E}+B$ & through &
  $g\times\mbox{2}^{E}+B+\mbox{2}^{k+D}-\mbox{1}$,\\
\end{displayLinesTable}
pages not occupied by an interrupt file should be read-only zeros.

@@

\begin{commentary}
Because the effect of a combined read-and-write of a \z{*topei} CSR
can be accomplished equally by a read of \z{*topei} followed by a write
of the interrupt identity to the corresponding \z{*clreipnum}, the
\z{*topei} CSRs could have been made read-only without the special
behavior defined for writes.
However, condensing these operations into a single CSR access intead of
two helps to reduce interrupt latency when critical.
\end{commentary}

>>>>

The layout of an interrupt-file's memory region is:\nopagebreak
\begin{displayLinesTable}[l@{\quad}l@{\qquad}l]
offset    & \ size  & register name \\
\noalign{\medskip}
\z{0x000} & 4 bytes & \z{seteipnum\_le} \\
\z{0x004} & 4 bytes & \z{seteipnum\_be} \\
\end{displayLinesTable}
All other bytes in an interrupt file's 4-KiB memory region are reserved
and must be implemented as read-only zeros.

@@

The requirements for read-only-zero pages apply only within each group.
Specifically, if $g$ is any integer between 0 and ${\mbox{2}^{j}-1}$
inclusive, then within the address ranges,
\begin{displayLinesTable}[lcll]
$g\times\mbox{2}^{E}+A$ & through &
  $g\times\mbox{2}^{E}+A+\mbox{2}^{k+C}-\mbox{1}$, &
  and\\
$g\times\mbox{2}^{E}+B$ & through &
  $g\times\mbox{2}^{E}+B+\mbox{2}^{k+D}-\mbox{1}$,\\
\end{displayLinesTable}
pages not occupied by an interrupt file should be read-only zeros.

@@

\begin{commentary}
Because the effect of a combined read-and-write of a \z{*topei} CSR
can be accomplished equally by a read of \z{*topei} followed by a write
of the interrupt identity to the corresponding \z{*clreipnum}, the
\z{*topei} CSRs could have been made read-only without the special
behavior defined for writes.
However, condensing these operations into a single CSR access instead of
two helps to reduce interrupt latency when critical.
\end{commentary}

### 2022-02-15 doc/src/IPIs.tex
Fix lots of typos

When harts have IMSICs, instead of using the Privileged Architecture's
mechanism for signalling software interrupts at remote harts, an IPI
can be sent to a hart by writing to the destination hart's IMSIC, the
same as a regular message-signalled interrupt (MSI)\@.
In that case, an incoming IPI appears at the destination hart as an
\emph{external interrupt} routed through the IMSIC, rather than as a
software interrupt as before.
However, so long as the same software (e.g.\@ an operating sytem or
machine monitor) is in control at both endpoints of an IPI, source
and destination, there should be no reason for a destination hart
to misinterpret the purpose of an incoming external interrupt that
represents an IPI.

>>>>

When harts have IMSICs, instead of using the Privileged Architecture's
mechanism for signalling software interrupts at remote harts, an IPI
can be sent to a hart by writing to the destination hart's IMSIC, the
same as a regular message-signalled interrupt (MSI)\@.
In that case, an incoming IPI appears at the destination hart as an
\emph{external interrupt} routed through the IMSIC, rather than as a
software interrupt as before.
However, so long as the same software (e.g.\@ an operating system or
machine monitor) is in control at both endpoints of an IPI, source
and destination, there should be no reason for a destination hart
to misinterpret the purpose of an incoming external interrupt that
represents an IPI.

### 2022-02-15 doc/src/VSLevel.tex
Fix lots of typos

When \z{hstatus}.VGEIN is the valid number of a guest interrupt
file, values of \z{vsiselect} in the range \z{0x70}--\z{0xFF} select
registers of this guest interrupt file, just as values of \z{siselect}
in the same range select registers of the IMSIC's true supervisor-level
interrupt file.
The registers of an interrupt file that are accessed indirectly through
\z{vsiselect} and \z{vireg} are documented in Chapter~\ref{ch:IMSIC}
on the IMSIC, along with the effects of accessing CSRs \z{vsseteipnum},
\z{vsclreipnum}, \z{vsseteienum}, \z{vsclreienum}, and \z{vstopei}.
Because all IMSIC interrupt files act identically, the guest interrupt
file that a virtual hart accesses through CSRs \z{siselect}, \z{sireg},
\z{sseteipnum}, \z{sclreipnum}, \z{sseteienum}, \z{sclreienum},
and \z{stopei} is indistinguishable from a true supervisor-level
interrupt file as seen from \mbox{S-mode} (or \mbox{HS-mode}).

@@

When bit VTI (Virtual Trap Interrupt control) =~1, attempts from
\mbox{VS-mode} to explictly access CSRs \z{sip}, \z{sie}, or, for RV32
only, \z{siph} or \z{sieh}, cause a virtual instruction exception.
Furthermore, when VTI =~1, a virtual instruction exception is raised
also for any attempt by the guest to write to a CSR that might
cause a bit of \z{vsip}/\z{vsiph} to change from one to zero,
excluding bit~9 for external interrupts (SEIP).
In particular, if register \z{vstimecmp} is implemented
(from extension Sstc), then attempts from \mbox{VS-mode}
to write to \z{stimecmp} or (RV32 only) \z{stimecmph}
cause a virtual instruction exception when VTI~=~1.

>>>>

When \z{hstatus}.VGEIN is the valid number of a guest interrupt
file, values of \z{vsiselect} in the range \z{0x70}--\z{0xFF} select
registers of this guest interrupt file, just as values of \z{siselect}
in the same range select registers of the IMSIC's true supervisor-level
interrupt file.
The registers of an interrupt file that are accessed indirectly through
\z{vsiselect} and \z{vsireg} are documented in Chapter~\ref{ch:IMSIC}
on the IMSIC, along with the effects of accessing CSRs \z{vsseteipnum},
\z{vsclreipnum}, \z{vsseteienum}, \z{vsclreienum}, and \z{vstopei}.
Because all IMSIC interrupt files act identically, the guest interrupt
file that a virtual hart accesses through CSRs \z{siselect}, \z{sireg},
\z{sseteipnum}, \z{sclreipnum}, \z{sseteienum}, \z{sclreienum},
and \z{stopei} is indistinguishable from a true supervisor-level
interrupt file as seen from \mbox{S-mode} (or \mbox{HS-mode}).

@@

When bit VTI (Virtual Trap Interrupt control) =~1, attempts from
\mbox{VS-mode} to explicitly access CSRs \z{sip}, \z{sie}, or, for RV32
only, \z{siph} or \z{sieh}, cause a virtual instruction exception.
Furthermore, when VTI =~1, a virtual instruction exception is raised
also for any attempt by the guest to write to a CSR that might
cause a bit of \z{vsip}/\z{vsiph} to change from one to zero,
excluding bit~9 for external interrupts (SEIP).
In particular, if register \z{vstimecmp} is implemented
(from extension Sstc), then attempts from \mbox{VS-mode}
to write to \z{stimecmp} or (RV32 only) \z{stimecmph}
cause a virtual instruction exception when VTI~=~1.

### 2022-02-15 doc/src/intro.tex
fix typo in intro.tex
Fix lots of typos

\item
Optionally allow software to configure the relative priorities of all
sources of interupts to a {\RISCV} hart (including the standard timer
and software interrupts, among others), instead of being limited
just to the ability of a separate interrupt controller to prioritize
external interrupts only.

@@

The Advanced Interrupt Architecture adds considerable support
for \emph{local interrupts} at a hart, whereby a hart essentially
interrupts itself in response to asynchronous events, usually errors.
Local interrupts remain contained within a hart (or close to it),
so like standard {\RISCV} timer and sofware interrupts, they do not
pass through an Advanced PLIC or IMSIC.

>>>>

\item
Optionally allow software to configure the relative priorities of all
sources of interrupts to a {\RISCV} hart (including the standard timer
and software interrupts, among others), instead of being limited
just to the ability of a separate interrupt controller to prioritize
external interrupts only.

@@

The Advanced Interrupt Architecture adds considerable support
for \emph{local interrupts} at a hart, whereby a hart essentially
interrupts itself in response to asynchronous events, usually errors.
Local interrupts remain contained within a hart (or close to it),
so like standard {\RISCV} timer and software interrupts, they do not
pass through an Advanced PLIC or IMSIC.
